---
title: "Alpha,Beta analysis"
author: "Adrià Auladell Martín"
output:
  pdf_document: 
    highlight: zenburn
  html_document: default
---

#Packages
```{r, message=FALSE, warning=FALSE}
library(vegan)
library(tidyverse)
library(phyloseq)
library(ggrepel)
library(patchwork)
library(scales)
library(lubridate)
library(lomb)
```

# Output of the script

```{r}

dir.create(file.path('../results/figures/timeseries'), showWarnings = FALSE)

```

#Data import

```{r}
bl.phy <- readRDS("../data/robjects/blphyloseq.rds") 

bl.phy <- filter_taxa(bl.phy, function(x) sum(x > 5) > 5, TRUE)

source("sourcefiles/params-graphs.R") 
source("sourcefiles/main_functions.R")

# Apart from the typical source of functions, we add the timeseries specific
source('sourcefiles/timeseries_functions.R')

physeq.list <- data_transformation(bl.phy)

bl.phy.relab <- physeq.list$relab
bl.phy.raref <- physeq.list$raref
bl.phy.log <- physeq.list$log
bl.phy.clr <- physeq.list$clr
bl.phy.asinh <- transform_sample_counts(bl.phy, function(x) asinh(x))

# since the list is quite big let's erase it
physeq.list <- NULL

tax <- as(tax_table(bl.phy), 'matrix') %>% 
  data.frame(asv = rownames(.)) %>% 
  mutate_if(.predicate = is.factor,.funs = as.character)

psmelt.relab <- psmelt_dplyr(bl.phy.relab)
psmelt.clr <- psmelt_dplyr(bl.phy.clr)
```


# Preprocessing

##Separate the two time series

Now we have to differentiate between the time series with both fractions, since we have to calculate the seasonality signal for each fraction separately. 

```{r}

bl.phy.0.2 <- subset_samples(bl.phy, fraction == '0.2') %>% 
  prune_taxa(taxa = taxa_sums(.) != 0, .)

bl.phy.3 <- subset_samples(bl.phy, fraction == '3') %>% 
  prune_taxa(taxa = taxa_sums(.) != 0, .)

# The rarefied counterparts
bl.phy.0.2.raref <- rarefy_even_depth(bl.phy.0.2,rngseed = 42)
bl.phy.3.raref <- rarefy_even_depth(bl.phy.3,rngseed = 42)

psm.bl.phy.0.2.raref <- psmelt_dplyr(bl.phy.0.2.raref)
psm.bl.phy.3.raref <- psmelt_dplyr(bl.phy.3.raref)
```

## Check completeness 

```{r}

# The 0.2 fraction presents all the data correctly

samdat.02 <- data.frame(sample_data(bl.phy.0.2)) 

samdat.02 %>% 
  .$year %>% 
  table()

samdat.02 %>% 
  .$month %>% 
  table()

samdat.3 <- data.frame(sample_data(bl.phy.3))

# There are some months lacking in the 3 fraction
samdat.3 %>% 
  .$year %>% 
  table()

table(samdat.3$year, samdat.3$month)
```

By looking at the data, we have the following samples unsampled:

- March 2004. 
- February 2005 , May 2005. 

We need to create a big df with all the information, and fill with NAs.

# Data preparation

## Filtering by occurrence 
```{r}
asvs.abundant02 <- readRDS('../data/robjects/asv_preval_02.rds') %>% 
  filter(behavior %in% c('Broad', 'Other')) %>% 
  pull(asv)

asvs.abundant3 <- readRDS('../data/robjects/asv_preval_3.rds') %>% 
  filter(behavior %in% c('Broad', 'Other')) %>% 
  pull(asv)


tsdf.0.2 <- transform_sample_counts(bl.phy.0.2, function(x) asinh(x)) %>% 
  prune_taxa(asvs.abundant02, . ) %>% 
  psmelt_dplyr() %>% 
  mutate(decimaldat = decimal_date(Date)) 

bl.phy.clr.02 <- subset_samples(bl.phy.clr, fraction == '0.2')

tsdf.0.2.clr <- bl.phy.clr.02 %>% 
  prune_taxa(asvs.abundant02, . ) %>% 
  psmelt_dplyr() %>% 
  mutate(decimaldat = decimal_date(Date)) 

tsdf.3 <- transform_sample_counts(bl.phy.3, function(x) asinh(x)) %>% 
  prune_taxa(asvs.abundant3, . ) %>% 
  psmelt_dplyr() %>% 
  mutate(decimaldat = decimal_date(Date))

bl.phy.clr.3 <- subset_samples(bl.phy.clr, fraction == '3')

tsdf.3.clr <- bl.phy.clr.3 %>% 
  prune_taxa(asvs.abundant3, . ) %>% 
  psmelt_dplyr() %>% 
  mutate(decimaldat = decimal_date(Date)) 



```



# Lombscargle 

### 0.2 fraction 

```{r}

df.ts <- tsdf.0.2.clr %>% 
  arrange(Date) %>% 
  select(Date, decimaldat, Abundance, OTU) %>% 
  split(.$OTU, drop = T)


  
lomb.02  <-  df.ts %>% 
  map(~randlsp( x =.x$Abundance,
                times = .x$decimaldat,
                type = 'period',
                plot = F))

lomb.sea.02 <- tibble( asv = names(lomb.02),
                       pval = map_dbl(lomb.02, ~.x$p.value),
                       peak = map_dbl(lomb.02, ~.x$peak),
                       interval  = map(lomb.02, ~.x$peak.at) )   %>% 
  filter(pval <= 0.01, peak >= 10)   %>% 
  pull(asv)


results.lomb02 <- lomb.02[lomb.sea.02]

map(results.lomb02, ~tibble( scanned = .x$scanned,
                          power = .x$power)) %>% 
  bind_rows(.id = 'asv') %>% 
  ggplot(aes(scanned, power)) + 
  geom_line(aes(group = asv)) + 
  facet_wrap(~asv) + 
  lil.strip
```

### 3 fraction 

```{r message=FALSE}

df.ts <- tsdf.3 %>% 
  arrange(Date) %>% 
  select(decimaldat, Abundance, OTU) %>% 
  split(.$OTU, drop = T)

  
lomb.3  <-  df.ts %>% 
  map(~randlsp( x =.x$Abundance,
                times = .x$decimaldat,
                type = 'period',
                plot = F))

lomb.sea.3 <- tibble( asv = names(lomb.3),
                      pval = map_dbl(lomb.3, ~.x$p.value), 
                      peak = map_dbl(lomb.3, ~.x$peak),
                      interval  = map(lomb.3, ~.x$peak.at) ) %>% 
  filter(pval <= 0.01, peak >= 10)   %>% 
  pull(asv)

results.lomb03 <- lomb.3[lomb.sea.3]

map(results.lomb03, ~tibble( scanned = .x$scanned,
                          power = .x$power)) %>% 
  bind_rows(.id = 'asv') %>% 
  ggplot(aes(scanned, power)) + 
  geom_line(aes(group = asv)) + 
  facet_wrap(~asv) + 
  theme_bw(base_size = 10) + 
  lil.strip

```

# Saving 

```{r}

saveRDS(lomb.02, "../data/robjects/results_ts_lomb02.rds")
saveRDS(lomb.3, "../data/robjects/results_ts_lomb3.rds")

```
