---
title: "Alpha,Beta analysis"
author: "Adrià Auladell Martín"
output:
  pdf_document: 
    highlight: zenburn
  html_document: default
---

```{r echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

#Packages
```{r message=FALSE, warning=FALSE}

set.seed(42) 

library(vegan)
library(tidyverse)
library(phyloseq)
library(ggrepel)
```

# Output of the script

```{r}


dir.create(file.path('../results/figures/betadiv'), showWarnings = FALSE)

```

#Data import

```{r}
bl.phy <- readRDS("../data/robjects/blphyloseq.rds") 
source("sourcefiles/params-graphs.R") 
source("sourcefiles/main_functions.R")

physeq.list <- data_transformation(bl.phy)

bl.phy.relab <- physeq.list$relab
bl.phy.rar <- physeq.list$raref
bl.phy.log <- physeq.list$log
bl.phy.clr <- physeq.list$clr

# since the list is quite big let's erase it
physeq.list <- NULL
```

## Separation of fractions 

```{r}
bl.phy.rar.02 <- subset_samples(bl.phy.rar,
                                                fraction == '0.2') %>% 
  prune_taxa(taxa = taxa_sums(.) > 0, x = .)

bl.phy.rar.3 <- subset_samples(bl.phy.rar,
                                                fraction == '3') %>% 
  prune_taxa(taxa = taxa_sums(.) > 0, x = .)


bl.phy.clr.02 <- subset_samples(bl.phy.clr,
                                fraction == '0.2') %>% 
  prune_taxa(taxa = taxa_sums(.) > 0, x = .)

bl.phy.clr.3 <- subset_samples(bl.phy.clr,
                               fraction == '3') %>% 
  prune_taxa(taxa = taxa_sums(.) > 0, x = .)

```

# Abundant and low 

```{r}
preval02 <- readRDS('../data/robjects/asv_preval_02.rds') 
preval3 <- readRDS('../data/robjects/asv_preval_3.rds') 

abund02 <- preval02 %>% filter(presence == 'abundant') %>% pull(asv)
abund3 <- preval3 %>% filter(presence == 'abundant') %>% pull(asv)
```


#Beta diversity estimation

Let's create the directory in which all results will be saved. 
```{r}

dir.create(file.path('../results/figures/betadiv'), showWarnings = FALSE)

```


##All 
### NMDS

```{r}

ordi.nmds.rare <- ordinate(bl.phy.rar, 'NMDS', 'bray')


nmds <- plot_ordination(bl.phy.rar,
                        ordi.nmds.rare, 
                        type='sample', justDF = TRUE) 

all <-  ggplot(nmds,aes(NMDS1,NMDS2)) +
  ggConvexHull::geom_convexhull(aes(group = fraction), alpha = 0.1) + 
  geom_point(aes(fill=season,shape=fraction), size = 3) +
  sea.fillScale + 
  coord_equal() +
  scale_y_continuous(n.breaks = 5) + 
  scale_shape_manual(values = c(21,24)) + 
  guides( fill = guide_legend(override.aes = list(shape = 21)))

ggsave("../results/figures/betadiv/general_allfrac_sams.pdf",
       all, 
       width =7,
       height = 5)

 
```

#### Only abundant
```{r}
bl.phy.rar.ab <- subset_taxa(bl.phy.rar,
                                (taxa_names(bl.phy.rar) %in% c(abund02,abund3)))

ordi.nmds.rare <- ordinate(bl.phy.rar.ab, 'NMDS', 'bray')


nmds <- plot_ordination(bl.phy.rar.ab,
                        ordi.nmds.rare, 
                        type='sample', justDF = TRUE) 

abonly <- ggplot(nmds,aes(NMDS1,NMDS2)) +
  ggConvexHull::geom_convexhull(aes(group = fraction), alpha = 0.1) + 
  # geom_line(aes(group = samname)) + 
  geom_point(aes(fill=season,shape=fraction), size = 3) +
  sea.fillScale + 
  coord_equal() +
  scale_y_continuous(n.breaks = 5) + 
  scale_shape_manual(values = c(21,24)) + 
  guides( fill = guide_legend(override.aes = list(shape = 21)))

ggsave("../results/figures/betadiv/general_allfrac_onlyab.pdf",
       abonly,
       width = 5, height =5)

samdat <-  data.frame(sample_data(bl.phy.rar.ab))

dist.ab <-  phyloseq::distance(bl.phy.rar.ab, method="bray")

season.adonis = adonis( dist.ab ~ season + fraction,
                        samdat)



```

#### Rare taxa only 

```{r}

bl.phy.rar.lowab <- subset_taxa(bl.phy.rar,
                                !(taxa_names(bl.phy.rar) %in% c(abund02,abund3)))

ordi.nmds.rare <- ordinate(bl.phy.rar.lowab, 'NMDS', 'bray')


nmds <- plot_ordination(bl.phy.rar.lowab,
                        ordi.nmds.rare, 
                        type='sample', justDF = TRUE) 

rarionly <- ggplot(nmds,aes(NMDS1,NMDS2)) +
  ggConvexHull::geom_convexhull(aes(group = fraction), alpha = 0.1) + 
  # geom_line(aes(group = samname), alpha = 0.4) +
  geom_point(aes(fill=season,shape=fraction), size = 3) +
  sea.fillScale + 
  coord_equal() +
  scale_y_continuous(n.breaks = 5) + 
  scale_shape_manual(values = c(21,24)) + 
  guides( fill = guide_legend(override.aes = list(shape = 21)))

ggsave("../results/figures/betadiv/general_allfrac_onlylowab.pdf",
       rarionly, 
       width = 5, height =5)

samdat <-  data.frame(sample_data(bl.phy.rar.lowab))

dist.rar <-  phyloseq::distance(bl.phy.rar.lowab, method="bray")
season.adonis = adonis( dist.rar ~ season + fraction,
                        samdat)

season.adonis
```

#### Composite 

```{r}
library(patchwork)

all_noguid <- all + guides(shape = FALSE,color = FALSE, fill = FALSE) 


composite <- (all_noguid | ((abonly + rarionly) / guide_area()) + 
  plot_layout(guides = 'collect',
              heights  = c(0.8, 0.2)) & leg.bottom) + 
  plot_annotation(tag_levels = 'A')

compov2 <- all + abonly + rarionly + 
  plot_annotation(tag_levels = 'A') + 
  plot_layout(guides = 'collect') &
  leg.bottom 

ggsave('../results/figures/betadiv/general_composite.pdf',
       plot = composite, 
       width = 12, height = 6)

ggsave('../results/figures/betadiv/general_compositev2.pdf',
       plot = compov2, 
       width = 12, height = 6)

```



#### Distances 

```{r}
  
put_distances_df <- function(physeq){
  
  data.frame(as.matrix(distance(physeq, method='bray'))) %>% 
  rownames_to_column(var = 'name.complete') %>% 
  gather(key = 'name.pair', value = 'bray', -name.complete)
}

comparison.distances <- c(bl.phy.rar.ab, bl.phy.rar.lowab) %>% 
  map(~put_distances_df(.x)) %>% 
  bind_rows( .id = 'presence') %>% 
  left_join(data.frame( sample_data(bl.phy)), by = 'name.complete') %>% 
  left_join(data.frame( sample_data(bl.phy)), by = c('name.pair' = 'name.complete')) %>% 
  filter(samname.x == samname.y) %>% 
  filter(name.complete != name.pair) %>% 
  mutate(presence = ifelse( presence == '1', 'abundant ASVs', 'rare ASVs'))
  


comparison.distances %>% 
  ggplot( aes(presence, 1 - bray)) + 
  geom_jitter(width = 0.1, aes(color = season.x)) + 
  geom_violin(width = 0.1, draw_quantiles = c(0.25,0.5,0.75), alpha = 0.9) + 
  sea.colScale + 
  ylab(' Bray curtis similarity') + 
  xlab(' Type of ASVs considered') 

ggsave('../results/figures/betadiv/distances_acrosssamedate.pdf',
       width = 6, height = 4)
```


## dbRDA
### Abstraction

### Adonis for selecting best variables 

```{r}
find_best_variables <- function(physeq, env, pval = 0.01){
  otu.adonis <- as(otu_table(physeq), "matrix") %>%
    data.frame() %>% 
    .[rownames(.) %in% rownames(env),] 
  print(dim(otu.adonis))
  print(dim(env))
    
  ad <- adonis(otu.adonis ~ ., data = env, permutations = 999, na.rm = T)
  print(ad)
  
  bestEnvVariables <- ad$aov.tab %>% 
    rownames() %>% 
    .[ad$aov.tab$"Pr(>F)" <= pval] %>% 
    .[!is.na(.)] 
  
  return(bestEnvVariables)
}


```

### Env dataset scaled

```{r}

env.df <- data.frame(sample_data(bl.phy)) %>% 
  rownames_to_column(var = "sam") %>% 
  select(sam,one_of("Temperature","Secchi", "Salinity", "Day_length",
                    "Chla_total", "PO4", "NH4", "NO2", "NO3", "Si", 
                    "PNF2_5um_Micro", "PNF_5um_Micro", "Cryptomonas",
                    "HNF_Micro", "HNA", "Prochlorococcus_FC", "Synechococcus")) %>% 
  na.omit() %>% 
  magrittr::set_rownames(.$sam)  %>% 
  select(-sam) %>% 
  scale() %>% 
  data.frame()
  

```

We will add too the pigment distribution values 

```{r}
odv_pigdata <- readxl::read_excel('../data/metadata-raw/Chemtax_final_ODV.xlsx')

odv.simplified <- odv_pigdata %>% 
  select(-c(1:7)) %>% 
  filter(Year %in% seq(2004,2009)) %>% 
  rename("year" = Year) %>% 
  mutate(month = str_pad(Month, width = 2, pad = '0')) %>% 
  # some dates are present more than once! 
  filter(! Data %in% c('23/03/2004', '24/03/2004',
                       '20/10/2004', '30/06/2009'))
  
  
  

scale_this <- function(x) as.vector(scale(x))

crossref <- get_variable(bl.phy, c('month', 'year')) %>% 
  rownames_to_column(var = 'sample') %>% 
  filter(sample %in% rownames(env.df)) %>% 
  left_join(odv.simplified, by = c('month', 'year')) %>% 
  mutate_if(is.double,scale_this) %>% 
  select(-month, -year, -Data, -Name, -Month, -Day)



env.df.pigments <- env.df %>% 
  rownames_to_column(var = 'sample') %>% 
  left_join(crossref, by = 'sample') %>% 
  filter(!is.na(PRAS)) %>% 
  column_to_rownames(var = 'sample') 

```


```{r}
GGally::ggcorr(env.df, label = T )
```


### Fraction 0.2 

```{r}


bestEnvVariables.fraction0.2 <- bl.phy.rar.02 %>% 
  find_best_variables(., env.df[sample_names(bl.phy.rar.02),])


dataset.fl <- distance_and_ordination(bl.phy.rar.02,
                                   bestEnvVariables.fraction0.2)


cleaning.vecdata <- dataset.fl$vectordata %>% 
  mutate( rowname = case_when(
    rowname == 'Day_length' ~ 'Day~length', 
    rowname == 'Chla_total' ~ 'Chl~italic(a)', 
    rowname == 'PNF_5um_Micro' ~ 'PNF~(5~mu*m)', 
    TRUE ~ rowname))

plot2 <- plot_dbrda(bl.phy.rar.02,
                   dataset.fl$ordination,
                   vector.data = cleaning.vecdata,
                   color = 'season') + 
  sea.colScale +
  leg.bottom + 
  scale_x_continuous(lim = c(-1.2, 1.5)) +
  scale_y_continuous(lim = c(-2, 1.4)) 

plot2 

ggsave(plot = plot2, '../results/figures/betadiv/dbrda_fraction2.pdf',
       height = 5, width = 5)

dataset$ordination
anova.cca(dataset$ordination, by="terms")
summary(eigenvals(dataset$ordination))[3,1:5]

find_best_variables(bl.phy.rar.02, env.df[sample_names(bl.phy.rar.02),] %>% 
                      select(bestEnvVariables.fraction0.2))

```


### Fraction 3 

```{r}

bestEnvVariables.fraction3 <- bl.phy.rar.3 %>% 
  find_best_variables(., env.df[sample_names(.),])

dataset <- distance_and_ordination(bl.phy.rar.3,
                                   bestEnvVariables.fraction3)


cleaning.vecdata <- dataset$vectordata %>% 
  mutate( rowname = case_when(
    rowname == 'Day_length' ~ 'Day~length', 
    rowname == 'Chla_total' ~ 'Chl~italic(a)', 
    rowname == 'PNF_5um_Micro' ~ 'PNF~(5~mu*m)', 
    rowname == 'PNF2_5um_Micro' ~ 'PNF~(2-5~mu*m)', 
    TRUE ~ rowname))


plot3 <- plot_dbrda(bl.phy.rar.3,
                   dataset$ordination,
                   vector.data = cleaning.vecdata,
                   color = 'season') + 
  sea.colScale + 
  leg.bottom + 
  scale_x_continuous(lim = c(-1.2, 1.5)) +
  scale_y_continuous(lim = c(-2, 1.4)) 

plot3

ggsave(plot = plot3, '../results/figures/betadiv/dbrda_fraction3.pdf',
       height = 5, width = 5)

```

## dbRDA composite 

```{r}
library(patchwork)
composite <- plot2 + plot3 + 
  plot_annotation(tag_levels = 'A') + 
  plot_layout(guides = 'collect') & 
  leg.bottom

ggsave(plot = composite, filename = '../results/figures/betadiv/dbrda_all.pdf',
       height =7, width = 10)

```

